---
title: " Murillo Alba Josué Iván 1. Datos, Modelos y Software. Ejercicios"
date: "Febrero 2024"
output: 
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---


<style>
body {
text-align: justify}
</style>

<style type="text/css">
body{/*Normal*/
     font-size: 20px;
}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
library(ggplot2)
library(plotly)
```



Invoke data set Mode from package mlogit. To do this you need to first load the package. This is a data set with choices about mode of transportation. This is done as follows:

```{r}
data("Mode")
```

Once you have loaded the data set, answer the following questions:
Describe this data set. How many variables are there and of which type (i.e., categorical/quantitative)?

```{r}
glimpse(Mode)
```


How many different modes of transportation are in this data set? What is the most popular mode? What is the least popular mode?

```{r}
table(Mode$choice)
summary(Mode$choice)
```

```{r}
ggplot(Mode)+ 
  geom_bar(aes(choice), color="black", fill="#6DB562")+
  theme_bw()
```

```{r}
summary(Mode)
```
In general, what is the most expensive mode? The least expensive?

```{r}
graf_costo<-ggplot(Mode) +
  geom_density(aes(cost.car, fill = "Costo de auto"), alpha = 0.6) +
  geom_density(aes(cost.carpool, fill = "Costo de carpools"), alpha = 0.6) +
  geom_density(aes(cost.bus, fill = "Costo de autobús"), alpha = 0.6) +
  geom_density(aes(cost.rail, fill = "Costo de ferrocarril"), alpha = 0.6) +
  scale_fill_manual(values = c("#0078BD", "#D61F33", "#F6D276", "#00B686"), name = "Variable") +
  labs(fill = "Variable") +
  theme_bw()

ggplotly(graf_costo)

```





Create a plot showing the univariate distributions of time by car and time by bus. Discuss.


```{r}
graf_tiempo<-ggplot(Mode) +
  geom_density(aes(time.car, fill = "Tiempo en auto"), alpha = 0.6) +
  geom_density(aes(time.carpool, fill = "Tiempo en carpools"), alpha = 0.6) +
  geom_density(aes(time.bus, fill = "Tiempo en autobús"), alpha = 0.6) +
  geom_density(aes(time.rail, fill = "Tiempo en ferrocarril"), alpha = 0.6) +
  scale_fill_manual(values = c("#0078BD", "#D61F33", "#F6D276", "#00B686"), name = "Variable") +
  labs(fill = "Variable") +
  theme_bw()

ggplotly(graf_tiempo)

```


How do choices relate to cost by the different modes?

```{r}
var_num<- names(Mode)[-1]

for (k in var_num) {
  graf<-ggplot(Mode)+
    geom_boxplot(aes(choice, eval(as.name(k)), fill=choice),
  show.legend= FALSE)+ ylab(k)+
    theme_bw()
  
  print(graf)
  
}
```




